Fractional Types

### 2019-11-17 fracGC/p.tex
a bit more on paper
abstract

\newcommand{\alt}{~|~}
\newcommand{\inlv}[1]{\ensuremath{\mathit{inl}(v)}}
\newcommand{\inrv}[1]{\ensuremath{\mathit{inr}(v)}}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\pointed}[2]{[#1 \bullet #2]}
\newcommand{\nboxtimes}[2]{\,\,~{^{#1}\boxtimes^{#2}}~\,\,}
\newcommand{\mm}{\texttt{\textminus}}
\newcommand{\pp}{\texttt{+}}
\newcommand{\inl}[1]{\textsf{inl}~#1}
\newcommand{\inr}[1]{\textsf{inr}~#1}
\newcommand{\idv}[3]{#2 \xrightarrow{#1} #3}
\newcommand{\cp}[3]{#1\stackrel{#2}{\bullet}#3}
\newcommand{\idt}[3]{#2 \equiv_{#1} #3}
\newcommand{\idrt}[3]{#3 \equiv_{#1} #2}
\newcommand{\refl}[1]{\textsf{refl}~#1}
\newcommand{\lid}{\textsf{lid}}
\newcommand{\rid}{\textsf{rid}}
\newcommand{\linv}{l!}
\newcommand{\rinv}{r!}
\newcommand{\invinv}{!!}
\newcommand{\assoc}{\circ}
\newcommand{\identlp}{\mathit{unite}_+\mathit{l}}
\newcommand{\identrp}{\mathit{uniti}_+\mathit{l}}
\newcommand{\identlsp}{\mathit{unite}_+\mathit{r}}
\newcommand{\identrsp}{\mathit{uniti}_+\mathit{r}}
\newcommand{\swapp}{\mathit{swap}_+}
\newcommand{\assoclp}{\mathit{assocl}_+}
\newcommand{\assocrp}{\mathit{assocr}_+}
\newcommand{\identlt}{\mathit{unite}_*\mathit{l}}
\newcommand{\identrt}{\mathit{uniti}_*\mathit{l}}
\newcommand{\identlst}{\mathit{unite}_*\mathit{r}}
\newcommand{\identrst}{\mathit{uniti}_*\mathit{r}}
\newcommand{\swapt}{\mathit{swap}_*}
\newcommand{\assoclt}{\mathit{assocl}_*}
\newcommand{\assocrt}{\mathit{assocr}_*}
\newcommand{\absorbr}{\mathit{absorbr}}
\newcommand{\absorbl}{\mathit{absorbl}}
\newcommand{\factorzr}{\mathit{factorzr}}
\newcommand{\factorzl}{\mathit{factorzl}}
\newcommand{\dist}{\mathit{dist}}
\newcommand{\factor}{\mathit{factor}}
\newcommand{\distl}{\mathit{distl}}
\newcommand{\factorl}{\mathit{factorl}}
\newcommand{\distz}{\mathit{absorbr}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\proves}{\vdash}
\newcommand{\idc}{\mathit{id}\!\!\leftrightarrow}
\newcommand{\Rule}[4]{
\makebox{{\rm #1}
$\displaystyle
\frac{\begin{array}{l}#2 \\\end{array}}
{\begin{array}{l}#3      \\\end{array}}$
 #4}}
\newcommand{\jdg}[3]{#2 \proves_{#1} #3}

@@

\begin{abstract}
Text of abstract \ldots.
\end{abstract}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Value of Ancillae}

@@

Def. of universe is:

\Udef{}

New combinators:

\EtaEpsilon{}

and their evaluation:

>>>>

\newcommand{\alt}{~|~}
\newcommand{\net}{t_\bullet}
\newcommand{\inlv}[1]{\ensuremath{\mathit{inl}(v)}}
\newcommand{\inrv}[1]{\ensuremath{\mathit{inr}(v)}}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\oneover}[1]{1/#1}
\newcommand{\pointed}[2]{[#1 \bullet #2]}
\newcommand{\nboxtimes}[2]{\,\,~{^{#1}\boxtimes^{#2}}~\,\,}
\newcommand{\mm}{\texttt{\textminus}}
\newcommand{\pp}{\texttt{+}}
\newcommand{\inl}[1]{\textsf{inl}~#1}
\newcommand{\inr}[1]{\textsf{inr}~#1}
\newcommand{\idv}[3]{#2 \xrightarrow{#1} #3}
\newcommand{\cp}[3]{#1\stackrel{#2}{\bullet}#3}
\newcommand{\idt}[3]{#2 \equiv_{#1} #3}
\newcommand{\idrt}[3]{#3 \equiv_{#1} #2}
\newcommand{\refl}[1]{\textsf{refl}~#1}
\newcommand{\lid}{\textsf{lid}}
\newcommand{\rid}{\textsf{rid}}
\newcommand{\linv}{l!}
\newcommand{\rinv}{r!}
\newcommand{\invinv}{!!}
\newcommand{\assoc}{\circ}
\newcommand{\identlp}{\mathit{unite}_+\mathit{l}}
\newcommand{\identrp}{\mathit{uniti}_+\mathit{l}}
\newcommand{\identlsp}{\mathit{unite}_+\mathit{r}}
\newcommand{\identrsp}{\mathit{uniti}_+\mathit{r}}
\newcommand{\swapp}{\mathit{swap}_+}
\newcommand{\assoclp}{\mathit{assocl}_+}
\newcommand{\assocrp}{\mathit{assocr}_+}
\newcommand{\identlt}{\mathit{unite}_*\mathit{l}}
\newcommand{\identrt}{\mathit{uniti}_*\mathit{l}}
\newcommand{\identlst}{\mathit{unite}_*\mathit{r}}
\newcommand{\identrst}{\mathit{uniti}_*\mathit{r}}
\newcommand{\swapt}{\mathit{swap}_*}
\newcommand{\assoclt}{\mathit{assocl}_*}
\newcommand{\assocrt}{\mathit{assocr}_*}
\newcommand{\absorbr}{\mathit{absorbr}}
\newcommand{\absorbl}{\mathit{absorbl}}
\newcommand{\factorzr}{\mathit{factorzr}}
\newcommand{\factorzl}{\mathit{factorzl}}
\newcommand{\dist}{\mathit{dist}}
\newcommand{\factor}{\mathit{factor}}
\newcommand{\distl}{\mathit{distl}}
\newcommand{\factorl}{\mathit{factorl}}
\newcommand{\distz}{\mathit{absorbr}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\proves}{\vdash}
\newcommand{\idc}{\mathit{id}\!\!\leftrightarrow}
\newcommand{\Rule}[4]{
\makebox{{\rm #1}
$\displaystyle
\frac{\begin{array}{l}#2 \\\end{array}}
{\begin{array}{l}#3      \\\end{array}}$
 #4}}
\newcommand{\jdg}[3]{#2 \proves_{#1} #3}

@@

\begin{abstract}
  In reversible and quantum computing, the management of space is
  subject to two broad classes of constraints. First, as is the case
  for general-purpose computation, every allocation must be paired
  with a matching de-allocation. Second, space can only be safely
  de-allocated if its contents are restored to their initial value
  from allocation time. Generally speaking, the state of the art
  provides limited partial solutions that address the first
  constraint by imposing a stack discipline and by leaving the second
  constraint to programmer's assertions.

  We propose a novel approach based on the idea of \emph{fractional
    types}. As a simple intuitive example, allocation of a new boolean
  value initialized to \textsf{false} also creates a value
  $\oneover{\textsf{false}}$ that can be thought of as a garbage
  collection (gc) process specialized to reclaim, and only reclaim,
  storage containing the value $\textsf{false}$. This gc process is a
  first-class entity that can sliced and diced by combining it with
  other gc processes and decomposing it in smaller
  processes.

  Technically, we formalize this idea in the context of a reversible
  language founded on type isomorphisms, prove its fundamental
  correctness properties, and illustrate its expressiveness using a
  wide variety of examples. The development is backed by a
  fully-formalized Agda implementation.
\end{abstract}

@@

General allocation and de-allocation. No need to keep track of
lifetime. Just make sure location is back to initial value. Safe to
de-allocate because indistinguishable from freshly created. No pointer
equality and aliases though. Concurrent: one thread allocates; gc
thread meets someone else

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Interchangeable Ancillae}

@@

We add a new type $\oneover{t}$ with a trivial runtime representation
$\ocircle$. We add two new level-1 programs to introduce and eliminate
the new type, and one new level-2 program that states that
introduction and elimination are inverses:
\[\begin{array}{lrcl}
\textit{Value types} & t &::=& \cdots \alt \oneover{t} \\
\textit{Non-empty types} & \net \\
\textit{Values}      & v &::=& \cdots \alt \ocircle \\
\textit{Level-1 types} &&& t \leftrightarrow t \\
\textit{Level-1 programs} & c &::=& \cdots \alt
   \eta : 1 \leftrightarrow (\net * \oneover{\net}) \\
   &&& ~~~~~\alt \epsilon : (\net * \oneover{\net}) \leftrightarrow 1 \\
\textit{Level-2 types} &&& c \Leftrightarrow c \\
\textit{Level-2 programs} & \alpha &::=& \cdots 
  \alt unity : \epsilon \circ \eta \Leftrightarrow \idc 
\end{array}\]
